# Disclosure risk assessment {#sec-6disclosure}

Once the microdata has been rendered anonymous or a conclusion has been made that it cannot be rendered anonymous without too much loss of utility, the Data Curator completes the disclosure risk assessment which includes information from the anonymization using SDC as well as complementary information on any other risks associated with disclosure, and writes the disclosure risk assessment (DRA) report. The Data Curator shares this report with the Data Provider, Data Controller and Data Protection Focal Point to inform the validation of the anonymization and authorization for the release of the anonymous version of the dataset on the MDL.

The DRA report must, at minimum, include the following information:

1.  Summary of the DRA

2.  List of UNHCR personnel, partners and/or other third parties performing each role (see Table 1) in the data curation process

3.  Summary of the potential disclosure risk scenarios

4.  Anonymization methods

<!-- -->

a.  Weights used for the anonymization

b.  List of variables removed from the dataset

c.  List of key variables used for the anonymization

d.  Statistical disclosure control (SDC) methods performed and on which variables

<!-- -->

5.  Assessment of the risk of re-identification of data subjects in the microdata

6.  Analysis of the utility of the anonymous microdata

7.  Conclusions

<!-- -->

a.  Can individual data subjects be identified by any means reasonably likely, based on the data alone or in combination with other data? \[Yes / No\]

b.  If necessary, are additional technical, organizational or legal measures or otherwise binding commitments in place to reduce the risk of disclosure? \[Yes / No\]

c.  If the microdataset will be published on the MDL, specify the proposed Terms of Use.

<!-- -->

8.  Any other relevant comments or explanations.

9.  Signatures, titles and dates (*Note: this is necessary for internal accountability for decision-making. As such, an email in lieu of signature is considered sufficient.*)

 

A [R Markdown](https://rmarkdown.rstudio.com/) DRA report template was developed to facilitate the linkage between the anonymization and the report. This is done by combining text and code chunks[^drr-1]. It is based off the UNHCR R Markdown report template ([`unhcrdown`](https://github.com/vidonne/unhcrdown)), which was developed to facilitate an easy-to-read report for both technical and non-technical audience. Before getting started, both need to be installed if not already.

[^drr-1]: For more information on R Markdown and code chunks: <https://rmarkdown.rstudio.com/lesson-3.html> and the full code chunk reference guide: <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf?_ga=2.24672569.2078611878.1633074546-1715259657.1633074546>

```{r, eval = F}

packages_needed = c("dplyr","foreign","ggplot2","haven","here",
                    "kableExtra","knitr", "labelled","lubridate",
                    "readr","readxl","rmarkdown","sampling", 
                    "sdcMicro","stringr","survey","tidyr","tidyverse",
                    "VIM", "writexl")
                    repos = c(CRAN = 'https://cran.rstudio.com', 
                              CRANextra = 'https://macos.rbind.io') 
                    install.packages(packages_needed[!packages_needed %in% row.names(installed.packages())], dependencies = TRUE, repos = repos)
                    update.packages(oldPkgs = packages_needed, ask = FALSE)
                    sapply(packages_needed, require, character.only=TRUE)

# UNHCR package for report template
remotes::install_github('vidonne/unhcrdown')
```

## R Markdown text formatting {#sec-6markdownformat}

The following table provides some of the syntax used in R Markdown for formatting text in the DRA report. Note this is only a sample of the various types of formatting that can be done in R Markdown. The following [cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.) is a useful tool to navigating all additional features:

|              Syntax               |       Description       |
|:---------------------------------:|:-----------------------:|
|            \# Header 1            |        Header 1         |
|           \## Header 2            |        Header 2         |
|           \### Header 3           |        Header 3         |
|           \### Header 4           |        Header 4         |
| \*\* bold \*\* and \_\_ bold \_\_ |  **bold** and **bold**  |
|  \* italics \* and \_ italics \_  | *italics* and *italics* |
|           \`code text\`           |       `code text`       |
|         \* unordered list         |   ·   unordered list    |
|            \+ sub-item            |        sub-item         |
|        1.     ordered list        |    1\. ordered list     |
|        2.     ordered list        |    2\. ordered list     |

 

## Getting started {#sec-6gettingstarted}

The template used below will call on the anonymization workspace and pull any data frames and other objects created. Ensure the workspace was saved prior to starting to write the report (see [Prepare weights](anonymize.qmd#sec-5weights)).

In cases of disproportionately stratified datasets, weights must be applied to analysis that spans across strata. The survey package can be used to apply the weights and run simple statistics.

```{r, eval = F}

# Create survey
your_dataset_survey <- svydesign(ids=~0, weights=~survey_weight,strata=~your_strata,survey.lonely.psu="adjust",data=your_dataset)
```

Now you can use your_dataset_survey to calculate the statistics needed. Note that it is more useful to compare proportions produced by the results as opposed to summaries because the summaries will be estimates for the entire population of interest in the sample frame not just within the sample

To start the report, open R studio and create a R Markdown file. Choose the HTML format and select the location to store the document.

Every R Markdown file starts with a YAML header, where the details of the document including title, date, author and template are specified. Details are always between - - -. The title and date will display on the first page of the report as well as in the header. It is defined from the beginning together with the report template.

*Creating the title and defining the template in R Markdown*

```{r, eval = F}

---
title: 'UNHCR Microdata Disclosure Risk Assessment Report'
author: 'First and Last Name - Position'
date: '`r Sys.Date()`’
output: unhcrdown::paged_simple
---

```

To be able to "call" on the anonymization script and pull any data frames and other objects created, the anonymization workspace needs to be loaded as well as the packages used. The following code is an example. Note that is it in a code chunk that will be hidden from the report.

About code chunks: Every code chunk should go between `{r}`. If the code output should be displayed but not the code itself, the option `{r, echo = FALSE}` is used. If neither the response nor the code should be displayed, the option `{r, include = FALSE}` should be used. More examples will be provided below.

*Setting up the workspace in R Markdown*

```{r, eval = F}

{r files, include=FALSE}
# Load the working space 
load( “1_scripts/your_workspace_name.RData"))
```

## The Report {#sec-6report}

### Introduction {#sec-6introduction}

The introduction should include the name of the dataset and the link to the dataset on RIDL. This can be written directly in R Markdown using, i.e. does not need to be part of a code chunk. Consider the following example.

```{r, eval = F}
# Introduction

__Dataset:__ "Côte d'Ivoire: Mapping of persons at risk of statelessness - April 2019"

__RIDL link:__ <https://ridl.unhcr.org/dataset/civ-statelessness-april-2019/>

```

### Summary {#sec-6summary}

The objective of the summary is to provide the main conclusions of the DRA. It should be one sentence or a short paragraph that includes the following information:

-   if microdata could be rendered anonymous while preserving their utility; and

-   if it is recommended to publish on the MDL and under which Terms of Use.

This can be written directly in R Markdown as a text editor:

```{r, eval = F}

# Summary

__These data could be anonymized while preserving their utility, and it is recommended that they be released as a licensed file on the Microdata Library.__
```

### List of roles {#sec-6roles}

The list of roles should include the names of everyone involved in the process including the Personal Data Controller, Data Provider(s), Data Curator and Data Protection Focal Point. This can be written directly in R Markdown using a table format for readability. Consider the following example.

```{r, eval = F}

Data Curation Roles                   | 
------------------------------------- | -------------
Personal Data Controller              | Name (email)
Data Provider                         | Name (email)
Data Curator                          | Name (email)

```

### Summary of disclosure risk scenarios

This section responds to question 1 in the DRA checklist: *What are the potential disclosure risk scenarios -- e.g., realistic assumptions about who may be interested in the microdata and for what purpose -- and available data and information (both internal and external) covering the same population group that could be linked to this personal microdata (e.g., through the mosaic effect)?*

It should be a short paragraph or two, and can be typed directly using R Markdown as a text editor. Consider the following example.

```{r, eval = F}
# Potential Disclosure Risk Scenarios

An intruder could be interested in identifying displaced individuals or households living in the camps or finding information about them that are observable such as specific household demographics, livelihood activities, living conditions, displacement experience and intentions for the immediate future.

The likelihood of re-identification of data subjects based on this dataset alone is low due to the fact that:
a) the anonymized dataset only includes a sample of the original full dataset,
b) observable variables were checked for granularity and recoded where the level of detail would increase risk of re-identification, and
c) the data was anonymized until 3-anonymity was reached.

The likelihood of re-identification of data subjects based on combining the anonymized version of the data with another publicly available dataset was not statistically measured, however is considered low due to the low level of publicly available on the data subjects.

```

### Anonymization methods

The anonymization methods section should include at minimum a description of the following:

-   weights: an example of a section of weights could look like this:

```{r}

## Weights
#Survey weights were calculated by dividing the total number of data subjects in the sample frame (N) by the number of data subjects in the sample (n) separately for each strata. The following table provides an overview:

strata <- c("Camp 1", "Camp 2", "Camp 3")
N <- c(500, 230, 475)
n <- c(140, 50, 125)
w <- c(3.5714,4.6000, 3.8000)
weight_table <- data.frame(strata,N,n,w)

colnames(weight_table) <- c("", "N", "n","weight")
library(kableExtra)
kbl(weight_table) %>%
  kable_classic()
```

 

-   variables removed

If any variables were removed from the dataset that will be released on the MDL, they should be listed here. If they were removed using a list (see example under [Remove variables that will not be published](anonymize.qmd#sec-5removeanon)), then it can simply be called and displayed. Consider the following example.

```{r, eval = F}
## Variable removed
The following *direct identifiers* were removed from the dataset:

print(direct_identifiers)
```

 

-   key variables

All key variables that were used as part of the SDC risk assessment should be listed here. If other variables were modified, but not used as part of the SDC risk assessment, they can be listed under a separate header (e.g. Variables modified). Consider the following example.

```{r, eval = F}
## Key variables
The following *key variables* were used for the SDC risk assessment:

  {r, echo=FALSE}
print(names(sdcObj@origData[,sdcObj@keyVars]))
```

 

-   SDC methods performed

Some of this information can be written directly in R Markdown as a text editors; others can be pulled in from the anonymization workspace as code chunks. Consider the following examples broken down by section.
